---
title: "insurance_EDA"
output: html_document
date: "2023-01-05"
author: Manasvi Ponaka
---

# Medical insurance cost prediction

```{r pressure, echo=FALSE}
library(tidyverse)
library(viridis)
library(gridExtra)
library(reshape2)
library(caret)
library(olsrr)
library(lmvar)
library(mctest)
library(groupdata2)
```

# Data Exploration and Manipulation

### Loading data

```{r ,echo=FALSE}
data_original <-read.csv("insurance.csv")
data <- data_original %>% mutate_if(is.character , as.factor)
```

```{r}
summary(data)
```

### Explore each column independently

```{r}
data %>% 
  ggplot(aes(x =age ))+
  geom_histogram(aes(y = ..density..),
                 bins = 30 , colour = 1, fill = "lightblue" , alpha = 0.5) +
  geom_density()+
  theme_classic()+
     labs(title= "Distribution of age across policy holders ")+
    theme(text=element_text(size=13,  family="Times") , plot.title = element_text( hjust = 0.5))+                
  theme(plot.title = element_text(hjust = 0.5, size=25),
              			  axis.title.x= element_text(size=25),
              			  axis.title.y= element_text(size=25),
              			  axis.text.x = element_text(size=25),
              			  axis.text.y = element_text(size=25),
              			  legend.title = element_text(size = 25),
              			  legend.text = element_text(size = 25),
              			  legend.position = "bottom")
```

```{r}
data %>% 
  group_by(sex) %>% 
  count() %>% 
  ggplot(aes( x = "" ,y = n , fill = sex))+
  geom_col( alpha= 0.5 , width=1, color="white")+
  geom_text(aes(label =n),position = position_stack(vjust = 0.5), color = "white", size=6)+
  coord_polar("y", start=0) + 
  theme_void() +
   labs(title= " Pi chart for gender profile of the policy holders ")+
    theme(text=element_text(size=13,  family="Times") , plot.title = element_text( hjust = 0.5))+
                  theme(plot.title = element_text(hjust = 0.2, size=25),
              			  legend.title = element_text(size = 25),
              			  legend.text = element_text(size = 25),
              			  legend.position = "right")

```

```{r}
data %>% 
  ggplot(aes(x =bmi) )+
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "lightblue" , alpha = 0.5) +
  geom_density()+
  theme_classic()+
   labs(title= "Distribution of bmi across policy holders ")+
    theme(text=element_text(size=13,  family="Times") , plot.title = element_text( hjust = 0.5))+
                theme(plot.title = element_text(hjust = 0.5, size=25),
              			  axis.title.x= element_text(size=25),
              			  axis.title.y= element_text(size=25),
              			  axis.text.x = element_text(size=25),
              			  axis.text.y = element_text(size=25),
              			  legend.title = element_text(size = 25),
              			  legend.text = element_text(size = 25),
              			  legend.position = "bottom")

```

```{r}
data %>% 
  mutate(across(c(children) ,as.factor)) %>% 
  group_by(children) %>% 
  count() %>% 
  ggplot(aes( x = "" ,y = n , fill = children))+
  geom_col( alpha= 0.5 , width=1, color="white")+
  geom_text(aes( x = 1.2 ,label = children),position = position_stack(vjust = 0.5), color = "white", size=4)+
  geom_text(aes(  x = 1.7, label =n),position = position_stack(vjust = 0.5), color = "black", size=3.5)+
  coord_polar("y", start=0) + 
  theme_void()+
  labs(title= " Pi chart of count of children for the policy holder")+
    theme(text=element_text(size=10,  family="Times") , plot.title = element_text( hjust = 0.5))+
                  theme(plot.title = element_text(hjust = 0.3, size=22),
              			  legend.title = element_text(size = 25),
              			  legend.text = element_text(size = 25),
              			  legend.position = "right")

```

```{r}
data %>% 
  mutate(across(c(children) ,as.factor)) %>% 
  group_by(children) %>% 
  count() %>% 
  ggplot(aes( x = children ,y = n , fill = children))+
  geom_col( alpha= 0.5 , width=1, color="white")+
  theme_classic()+
     labs(title= "Barplot of children count across policy holders \n")+
  theme(text=element_text(size=13,  family="Times") , plot.title = element_text( hjust = 0.5))+
                  theme(plot.title = element_text(hjust = 0.4, size=25),
              			  axis.title.x= element_text(size=25),
              			  axis.title.y= element_text(size=25),
              			  axis.text.x = element_text(size=25),
              			  axis.text.y = element_text(size=25),
              			  legend.title = element_text(size = 25),
              			  legend.text = element_text(size = 25),
              			  legend.position = "right")

```

```{r}
data %>% 
  mutate(across(c(smoker) ,as.factor)) %>% 
  group_by(smoker) %>% 
  count() %>% 
  ggplot(aes( x = "" ,y = n , fill = smoker))+
  geom_col( alpha= 0.5 , width=1, color="white")+
  geom_text(aes( x = 1.2 ,label = smoker),position = position_stack(vjust = 0.5), color = "white", size=7)+
  geom_text(aes(  x = 1.7, label =n),position = position_stack(vjust = 0.5), color = "black", size=7)+
  coord_polar("y", start=0) + 
  theme_void()+
  labs(title= " Pi chart of count of children for the policy holder ")+
    theme(text=element_text(size=13,  family="Times") , plot.title = element_text( hjust = 0.5))+                
  theme(plot.title = element_text(hjust = 0.3, size=24),
              			  legend.title = element_text(size = 25),
              			  legend.text = element_text(size = 25),
              			  legend.position = "right")
```

```{r}
data %>% 
  mutate(across(c(region) ,as.factor)) %>% 
  group_by(region) %>% 
  count() %>% 
  ggplot(aes( x = region ,y = n , fill = region))+
  geom_col( alpha= 0.5 , width=1, color="white")+
  theme_classic()+
     labs(title= "Barplot of regions across policy holders ")+
  theme(text=element_text(size=13,  family="Times") , plot.title = element_text( hjust = 0.5))+                
  theme(plot.title = element_text(hjust = 0.5, size=24),
              			  axis.title.x= element_text(size=0),
              			  axis.title.y= element_text(size=25),
              			  axis.text.x = element_text(size=0 , angle = 45),
              			  axis.text.y = element_text(size=25),
              			  legend.title = element_text(size = 20),
              			  legend.text = element_text(size = 20),
              			  legend.position = "right")
```

```{r}
data %>% 
  ggplot(aes(x = charges) )+
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "lightblue" , alpha = 0.5) +
  geom_density()+
  theme_classic()+
   labs(title= "Distribution of charges across policy holders ")+
    theme(text=element_text(size=13,  family="Times") , plot.title = element_text( hjust = 0.8))+                
  theme(plot.title = element_text(hjust = 0.9, size=25),
              			  axis.title.x= element_text(size=25),
              			  axis.title.y= element_text(size=25),
              			  axis.text.x = element_text(size=25),
              			  axis.text.y = element_text(size=25),
              			  legend.title = element_text(size = 25),
              			  legend.text = element_text(size = 25),
              			  legend.position = "bottom")
```

### Interacting variable analysis

```{r}
data %>% 
    mutate(across(c(smoker) ,as.factor)) %>%
  group_by(smoker) %>%
  ggplot(aes( x= age , y = charges , colour = smoker ))+
  geom_point(alpha= 0.5)+
  geom_smooth(method = "lm" , se = FALSE)+
    theme_classic()+
  labs(title= "Point plot for age vs charges across smoking and \n non smoking policy holders")+
  theme(text=element_text(size=10,  family="Times") , plot.title = element_text( hjust = 0.5))+
                  theme(plot.title = element_text(hjust = 0.5, size=20),
              			  axis.title.x= element_text(size=20),
              			  axis.title.y= element_text(size=20),
              			  axis.text.x = element_text(size=20),
              			  axis.text.y = element_text(size = 20),
              			  legend.title = element_text(size = 20),
              			  legend.text = element_text(size = 20),
              			  legend.position = "right")
```

```{r}
data %>% 
    mutate(across(c(smoker) ,as.factor)) %>%
  group_by(smoker) %>%
  ggplot(aes( x= bmi , y = charges , colour = smoker))+
  geom_point(alpha= 0.5)+
  geom_smooth(method = "lm" , se = FALSE)+
    theme_classic()+
  labs(title= "Point plot for charges vs bmi across smoking and \n non smoking policy holders")+
  theme(text=element_text(size=10,  family="Times") , plot.title = element_text( hjust = 0.5))+
                  theme(plot.title = element_text(hjust = 0.5, size=20),
              			  axis.title.x= element_text(size=20),
              			  axis.title.y= element_text(size=20),
              			  axis.text.x = element_text(size=20),
              			  axis.text.y = element_text(size = 20),
              			  legend.title = element_text(size = 20),
              			  legend.text = element_text(size = 20),
              			  legend.position = "right")
```

```{r}
box1<-data %>% 
  group_by(smoker) %>%
  ggplot(aes( x= region , y = charges , fill = region))+
  geom_boxplot(alpha= 0.5)+
    theme_classic()+
  labs(title= "Box plot for charges across policy holders of different regions ")+
  theme(text=element_text(size=10,  family="Times") , plot.title = element_text( hjust = 0.5))
box1+ 
  theme(text=element_text(size=10,  family="Times") , plot.title = element_text( hjust = 0.5))+
                  theme(plot.title = element_text(hjust = 0.335, size=20),
              			  axis.title.x= element_text(size=20),
              			  axis.title.y= element_text(size=20),
              			  axis.text.x = element_text(angle = 30, hjust  = 0.9,size=20),
              			  axis.text.y = element_text(size = 20),
              			  legend.title = element_text(size = 20),
              			  legend.text = element_text(size = 20),
              			  legend.position = "right")
```

```{r}
box2 <-data %>% 
  mutate(across(c(children) ,as.factor)) %>%
  group_by(children) %>%
  ggplot(aes( x= children , y = charges , fill = children))+
  geom_boxplot(alpha= 0.5)+
    theme_classic()+
  labs(title= "Box plot for charges across policy holders with \n different children count ")+
  theme(text=element_text( family="Times") , plot.title = element_text( hjust = 0.5))+
                  theme(plot.title = element_text(hjust = 0.5, size=20),
              			  axis.title.x= element_text(size=20),
              			  axis.title.y= element_text(size=20),
              			  axis.text.x = element_text(size=20),
              			  axis.text.y = element_text(size = 20),
              			  legend.title = element_text(size = 20),
              			  legend.text = element_text(size = 20),
              			  legend.position = "right")
box2  

```

```{r}
box3<-data %>%
  group_by(sex) %>%
  ggplot(aes( x= sex , y = charges , fill = sex))+
  geom_boxplot(alpha= 0.5)+
    theme_classic()+
  labs(title= "Box plot for charges across policy holders of different regions ")+
  theme(text=element_text(size=10,  family="Times") , plot.title = element_text( hjust = 0.5))
box3+ 
                  theme(plot.title = element_text(hjust = 0.4, size=20),
              			  axis.title.x= element_text(size=20),
              			  axis.title.y= element_text(size=20),
              			  axis.text.x = element_text(size=20),
              			  axis.text.y = element_text(size = 20),
              			  legend.title = element_text(size = 20),
              			  legend.text = element_text(size = 20),
              			  legend.position = "right")
```

### Data Pre-processing

```{r}

# anonymus function to generate binary columns
minus1<-function(x){
  x-1
}

#change categorical data to numerical data
data <- data %>%
  mutate(across(where(is.factor) & !region,as.double)) %>% 
  mutate(across(where(is.double),minus1)) %>% 
  mutate(region_northeast = ifelse(region=='northeast', 1, 0),
         region_northwest = ifelse(region=='northwest', 1, 0),
         region_southeast = ifelse(region=='southeast', 1, 0),
         region_southwest = ifelse(region=='southwest', 1, 0)) %>% 
  select(-region)


#generate polynomial features for numerical data
poly_feat <- data %>% 
             select(!charges) %>% 
             mutate(exposure = log(data$age) , age2 =  age^2 ,age3 =  age^3, bmi2 = bmi^2 ,bmi3 = bmi^3,age_bmi_2 = age*bmi )

head(poly_feat)
```

```{r}
cormat <-cor(data %>% mutate(charges = data$charges))

# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
  upper_tri <- get_upper_tri(round(cormat ,2))

  
  # Melt the correlation matrix
library(reshape2)
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Heatmap
library(ggplot2)
ggheatmap <- ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_void()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 8, hjust = 1) , axis.text.y = element_text(vjust = 1, 
    size = 8, hjust = 1))+
 coord_fixed()

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 2) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```

### splitting data to train and test

```{r}

# function to split dataframe into test and train according to given partition size 
gen_train_test<- function(data , partition_size , normalise_flag ){
  set.seed(123)
  training.samples <- data$charges %>%
  createDataPartition(p = partition_size, list = FALSE)

  train.data  <- data[training.samples, ]
  test.data <- data[-training.samples, ]

  if (normalise_flag){
  normalizer <- preProcess(train.data, method="range")
  train.data <-predict(normalizer , train.data)
  test.data <-predict(normalizer , test.data)}


  train.data$charges <- data$charges[training.samples]
  test.data$charges <- data$charges[-training.samples]

return(list(train.data , test.data))

}


```

```{r}
temp_data <- gen_train_test(data , partition_size = 0.8 , TRUE)
  train.data <- temp_data[[1]]
  test.data <- temp_data[[2]]

data_original<- data_original %>% mutate_if(is.character , as.factor)
data_original$children <- as.factor(data_original$children)
temp_data <- gen_train_test(data_original , partition_size = 0.8 ,FALSE)
  train.data.original <- temp_data[[1]]
  test.data.original <- temp_data[[2]]
```

# Multi-linear Model

## Charges \~ all columns excluding bmi

```{r}
# Fitting charges ~ all columns excluding bmi
sm1 <- lm(log(charges)~ age+sex+children+smoker+region, data = train.data.original)

# Take a look on summary of the model
summary(sm1)

pred_sm1 <- predict(sm1 , test.data.original)

sqrt(mean((exp(pred_sm1) - test.data.original$charges) ^ 2))
```

From the above result we can say that the model is statistically significant having p-value \< 2.2e-16 and coefficients are relevant.

## Charges \~ all columns excluding age

```{r}
# Fitting charges ~ all columns excluding age
sm2 <- lm(log(charges) ~ bmi+sex+children+smoker+region, data = train.data.original)

# Take a look on summary of the model
summary(sm2)

pred_sm2 <- predict(sm2 , test.data.original)

sqrt(mean((exp(pred_sm2) - test.data.original$charges) ^ 2))
```

From the above result we can say that the model is statistically significant having p-value \< 2.2e-16 and coefficients are relevant.

## Charges \~ all columns - Baseline model

```{r}
# Fitting charges ~ all columns 
sm3 <- lm(log(charges) ~ age +bmi+sex+children+smoker+region, data = train.data.original)

# Take a look on summary of the model
summary(sm3)

```

#### metrics function

```{r}
metrics<- function(pred , data , log_flag = TRUE, train_flag = TRUE){
  
  if (log_flag){
  pred <- as.numeric(unlist(lapply(pred, exp)))
  }
  
  mae <- mean(abs(pred- data$charges))

  rmse<- sqrt(mean((pred - data$charges) ^ 2))

  r2<- 1 - (sum((pred - data$charges)^2) / sum((pred - mean(data$charges))^2))

 if(train_flag){cat("Train\n")}else{  cat("Test\n")}

  cat("\n rmse :" , rmse, "\n r2:", r2 ,"\n",'mae: ' , mae ,"\n")
  return(c(rmse , r2 ,mae))
}
```

```{r}

train_pred_sm3 <- predict(sm3 , train.data.original)

metrics(train_pred_sm3 , train.data.original)

pred_sm3 <- predict(sm3 , test.data.original )

metrics(pred_sm3 , test.data.original,train_flag=  FALSE)


```

```{r}

par(mfrow = c(2, 2))

plot(sm3)
mtext("Residual analysis plot for simple linear model\n\n", side = 3, line = - 2, outer = TRUE)

```

#### Residual Normality test

```{r}
hist(sm3$residuals)
```

> `The residual qqplot is deviating from 45 degree indicating that the residuals are not normal and its supported by the skewed histogram`

Lets test for normality

```{r}
shapiro.test(sm3$residuals)
```

\>`p-value < 0.05 , Rejecting the null hypothesis that the residuals are normal`

#### Heteroscedascity test

```{r}
olsrr::ols_test_score(sm3)

```

\>`The p-value < 0.05 hence rejecting the null hypothesis that the variance is homogeneous`

#### Multi collinearity

```{r}

omcdiag(sm3)
```

```{r}
imcdiag(sm3)
```

\>`No Multicollinearity in the data`

\>`This tests indicate that there might presence of unoticeable outliers in the data and/or the relation ship between the dependent and independent variables in the data is not linear. We can see the hints of this from our interaction plots in the previous sections.`

\>`Let's first proceed with the assumption that the data is not linear and start with a polynomial regresion and see if its enough to fit our complex data .`

# Polynomial Regression

```{r}
pm1 <- lm(log(charges) ~ poly(age , 2)+poly(age , 2)+children+smoker+region  , data = train.data.original)

summary(pm1)
```

From the above result we can say that the model is statistically significant having p-value \< 2.2e-16 and coefficients are relevant.

```{r}
pred_pm1 <- predict(pm1 , train.data.original)
metrics(pred_pm1 ,train.data.original , train_flag = TRUE)

pred_pm1 <- predict(pm1 , test.data.original)
metrics(pred_pm1 ,test.data.original , train_flag = FALSE)
```

\>`Lets check fo other models before introducing an interaction term to model.`

# Other models

```{r}
# Run algorithms using 10-fold cross validation
control <- trainControl(method="cv", number=10)
metric <- "RMSE"
```

## a) nonlinear algorithms - CART , KNN

```{r}

# CART
set.seed(7)
fit.cart <- train(charges~., data=train.data.original, method="rpart", metric=metric, trControl=control)

print(fit.cart)

train.pred_cart <- predict(fit.cart , data = train.data.original)

metrics(train.pred_cart , train.data.original , log_flag = FALSE )

pred_cart <- predict(fit.cart , newdata = test.data.original )

metrics(pred_cart , test.data.original, log_flag = FALSE , train_flag = FALSE)

# kNN
set.seed(7)
fit.knn <- train(charges~., data=train.data.original, method="knn", metric=metric, trControl=control)

print(fit.knn)

train.pred_knn <- predict(fit.knn , data = train.data.original)

metrics(train.pred_knn , train.data.original, log_flag = FALSE)

pred_knn <- predict(fit.knn , newdata = test.data.original )

metrics(pred_knn , test.data.original , log_flag = FALSE , train_flag = FALSE)
```

## b) advanced algorithms - SVM , Random Forest

```{r}
# b) advanced algorithms
# SVM
set.seed(7)
fit.svm <- train(charges~., data=train.data.original, method="svmRadial", metric=metric, trControl=control)

print(fit.svm)

train.pred_svm <- predict(fit.svm , data = train.data.original)

metrics(train.pred_svm , train.data.original, log_flag = FALSE)

pred_svm <- predict(fit.svm , newdata = test.data.original )

metrics(pred_svm , test.data.original , log_flag = FALSE , train_flag = FALSE)



# Random Forest
set.seed(7)
fit.rf <- train(charges~., data=train.data.original, method="rf", metric=metric, trControl=control)

print(fit.rf)

train.pred_rf<- predict(fit.rf , data = train.data.original)

metrics(train.pred_rf, train.data.original, log_flag = FALSE)

pred_rf <- predict(fit.rf , newdata = test.data.original )

metrics(pred_rf , test.data.original , log_flag = FALSE , train_flag = FALSE)
```

## Model with interactions

```{r}

### Function to return trained model
lmi_out <- function(train, test) {
model.lmi<- lm(log(charges)~age+sex+bmi+children+smoker+region+age:smoker+bmi:smoker+I(log(age))+I(log(bmi)),data = train , x = TRUE ,y = TRUE)

print(summary(model.lmi))

train.pred_lmi <- predict(model.lmi , train)

metrics(train.pred_lmi , train)

pred_lmi <- predict(model.lmi , newdata = test )

metrics(pred_lmi , test ,train_flag=  FALSE)

return(model.lmi)
}
```

### Cross validation

```{r}
#Function to print output scores of the model

lmi_out.metrics<- function(train, test , fold) {
model.lmi <- lm(log(charges)~age+sex+bmi+children+smoker+region+age:smoker+bmi:smoker+I(log(age))+I(log(bmi)),data = train , x = TRUE ,y = TRUE)

print(summary(model.lmi))

pred_lmi <- predict(model.lmi , test)

rmse <- sqrt(mean((exp(pred_lmi) - test$charges) ^ 2))

mae <- mean(abs(exp(pred_lmi) - test$charges))

r2<- 1 - (sum((exp(pred_lmi) - test$charges)^2)) / (sum((exp(pred_lmi) - mean(test$charges))^2))

cat(fold , "\n" , "rmse: ", "r2:" , r2,rmse , "mae: ",mae )

return(c(rmse ,r2, mae))
}
```

```{r}
set.seed(7)
data_cv <- fold(
  train.data.original,
  k = 10,
  cat_col = c("children","smoker" , "sex" ,"region"),
  num_fold_cols = 10,
  parallel = FALSE # set to TRUE to run in parallel
)


n <- 10 ##number of folds in cv
scores <- matrix(ncol=3, nrow=n)


for (i in seq(1, 10)){
   
  fold <- paste(".folds_" , toString(i) , sep = "")
  
  test_set <- data_cv %>% 
    dplyr::filter(!!as.symbol(fold) == 3)
  
  train_set <- data_cv %>% 
    dplyr::filter(!!as.symbol(fold) != 3)

  scores[i ,]<- lmi_out.metrics(train_set , train_set , fold)
}
```

```{r}
# column1 <- rmse , column2 <- mae
mean_rmse <- mean(scores[,1])
mean_mae <- mean(scores[,3])
mean_r2 <- mean(scores[,2])

cat(" mean rmse of the model: " , mean_rmse ,"\n", "mean mae of the model:" ,  mean_mae,"\n", "mean r-squared of the model:" ,  mean_r2)
```

```{r}
model.lmi1 <- lmi_out(train.data.original , test.data.original)
```

```{r}
par(mfrow = c(2, 2))
plot(model.lmi1 , main = "Non-Linear model residual analysis plot\n")

```

\>\`The model

### outliers

lets delete the outliers using cook's method fit the model

#### cooks - Not recommended as it does not satisfy linearity assumptions

```{r}

cooksD <- cooks.distance(model.lmi1)
influential <- cooksD[(cooksD > (3 * mean(cooksD, na.rm = TRUE)))]
outliers <- data_original[names(influential),]
data_no_out <- data_original %>% anti_join(outliers)

```

The assumptions (linear relation ship of dependent and independent variables) of the cooks method are not met by the data.So, this method id nort recommended.

#### mahalanobisb- recommended

```{r}

# Finding the center point 
data.center  = colMeans(data_original %>% select(where(is.numeric)))

# Finding the covariance matrix
data.cov     = cov(data_original %>% select(where(is.numeric)))


# Finding distances
distances <- mahalanobis(x = data_original %>% select(where(is.numeric)) , center = data.center , cov = data.cov)

# Cutoff value for distances from Chi-Square Dist. 
# with p = 0.95 df = 2 which in ncol(air)
cutoff <- qchisq(p = 0.95 , df = ncol(data_original %>% select(where(is.numeric))))

## outliers observation whose distance greater than cutoff value
influential.<- row.names(data_original[distances < cutoff ,])

outliers <- data_original[names(influential),]

data_no_out <- data_original %>% anti_join(outliers)

```

```{r}
set.seed(7)
temp_data <- gen_train_test(data_no_out , partition_size = 0.8 ,FALSE)
  train.data.no_out <- temp_data[[1]]
  test.data.no_out <- temp_data[[2]]
```

### Cross validation

```{r}

set.seed(7)
data_cv_out <- fold(
  train.data.no_out,
  k = 10,
  cat_col = c("children","smoker" , "sex" ,"region"),
  num_fold_cols = 10,
  parallel = FALSE # set to TRUE to run in parallel
)



n <- 10
scores_out <- matrix(ncol=3, nrow=n)


for (i in seq(1, 10)){
   
  fold <- paste(".folds_" , toString(i) , sep = "")
  
  test_set <- data_cv_out %>% 
    dplyr::filter(!!as.symbol(fold) == 3)
  
  train_set <- data_cv_out %>% 
    dplyr::filter(!!as.symbol(fold) != 3)

  scores_out[i ,]<- lmi_out.metrics(train_set , train_set , fold)
}
```

```{r}
# column1 <- rmse , column2 <- mae
mean_rmse <- mean(scores_out[,1])
mean_mae <- mean(scores_out[,3])
mean_r2 <- mean(scores_out[,2])

cat(" mean rmse of the model: " , mean_rmse ,"\n", "mean mae of the model:" ,  mean_mae,"\n", "mean r-squared of the model:" ,  mean_r2)
```

```{r}
model.lmi.no_out <- lmi_out(train.data.no_out ,test.data.no_out )
par(mfrow = c(2, 2))
plot(model.lmi.no_out)
```

```{r}
olsrr::ols_test_score(model.lmi.no_out)
```
We reject the alternative hypothesis that variamnce is not homogenous.

```{r}
omcdiag(model.lmi.no_out)
```

```{r}
imcdiag(model.lmi.no_out)

```

### Residual Analysis

```{r}
par(mfrow = c(2, 2))

plot(model.lmi.no_out)
resid_hist<- ggplot(model.lmi.no_out, aes( x = .resid)) +
  geom_histogram(aes(y = ..density.. ), color = "black" , fill = "lightblue") +
  geom_density(fill = "red", alpha = 0.2)+
  labs(title= "Histogram of residuals")+
  theme_classic()+
  theme(text=element_text(size=10,  family="Times") , plot.title = element_text( hjust = 0.5))+
                  theme(plot.title = element_text(hjust = 0.4, size=20),
              			  axis.title.x= element_text(size=20),
              			  axis.title.y= element_text(size=20),
              			  axis.text.x = element_text(size=20),
              			  axis.text.y = element_text(size = 20),
              			  legend.title = element_text(size = 20),
              			  legend.text = element_text(size = 20),
              			  legend.position = "right")


resid_hist
```
>`Residuals are randomly distributed and their variance is homogeneous as confirmed by ols_test`

```{r}
hist(train.data.no_out$charges,col = "lightblue",probability = TRUE,lwd=2,breaks = 35 ,main = "Distribution of Train data Predicted and True value")
lines(density(train.data.no_out$charges),col="red",lty=2,lwd=3)

lines( density(exp(model.lmi.no_out$fitted.values)),col="blue",lty=2,lwd=3)

legend("topright",legend = c("charges","exp(model fitted values)"),col = c("red","blue"),lty = 2,lwd=2)
```
```{r}
pred_nlm <- predict(model.lmi.no_out , test.data.no_out)

hist(test.data.no_out$charges,col = "lightblue",breaks = 35,probability = TRUE , main = "Distribution of Test data Predicted and True value")
lines(density(test.data.no_out$charges),col="red",lwd=3,lty=2)
lines(density(exp(pred_nlm)),col="blue",lty=2,lwd=3)
legend("topright",legend = c("charges","exp(model fitted values)"),col = c("red","blue"),lty = 2,lwd=2)

```
